---
title: "Bias-corrected Bayesian meta-analysis for TAAR1 agonists"
subtitle: "Combining animal and human evidence using the Verde (2021) jarbes framework"
author: "Alexandra Bannach-Brown"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  cache   = FALSE
)

# Helper: one-row posterior summary data frame (bind multiple for a table)
post_summary <- function(samps, label) {
  data.frame(
    Parameter        = label,
    `Mean [95% CrI]` = sprintf("%.3f [%.3f, %.3f]",
                               mean(samps),
                               quantile(samps, 0.025),
                               quantile(samps, 0.975)),
    check.names = FALSE
  )
}
```

---

# Purpose

This document demonstrates a **bias-corrected Bayesian meta-analysis** that
formally combines animal and human evidence on TAAR1 agonists for
schizophrenia. Rather than analysing the two evidence streams independently and
comparing results informally, the approach introduced by Verde (2021) treats
the animal studies as potentially biased observations and uses the human RCTs
as an anchor to estimate and correct for that bias.

Five models are fitted and compared:

| # | Model | Prior on bias fraction |
|---|-------|------------------------|
| 1 | **RCT-only RE** | — (single evidence stream) |
| 2 | **Animal-only RE** | — (single evidence stream) |
| 3 | **BC (informative)** | Beta derived from study counts |
| 4 | **BC (non-informative)** | Beta(1, 1) = Uniform |
| 5 | **PBias (ROB covariate)** | Logistic regression on ROB score |

**Key references**

- Verde PE (2021). *A bias-corrected meta-analysis model for combining studies
  of different types and quality.* Biometrical Journal 63(2):406–422.
  <https://doi.org/10.1002/bimj.201900376>
- Siafis et al (2024). GALENOS TAAR1 living systematic review.
  <https://pmc.ncbi.nlm.nih.gov/articles/PMC11258611/>
- Smith et al (2025). GALENOS TAAR1 triangulation paper.
  <https://www.cambridge.org/core/journals/the-british-journal-of-psychiatry/article/triangulating-evidence-from-the-galenos-living-systematic-review-on-trace-amineassociated-receptor-1-taar1-agonists-in-psychosis/B5202B48429F1E700F5274724FB176C0>

---

# Statistical framework

## The Verde (2021) bias-correction idea

The core intuition is that animal and human studies measure a related but not
identical quantity. Animal studies are often inflated by several sources of
systematic bias (publication pressure, lack of blinding, translational gaps)
so their pooled effect $\mu_{\text{animal}}$ tends to exceed the clinical
effect $\mu_{\text{RCT}}$ by a constant amount $B > 0$.

Verde's **bias-corrected (BC) model** captures this with a two-component
mixture:

$$y_i \sim N(\theta_i^{\text{BC}},\; \sigma_i^2)$$

$$\theta_i^{\text{BC}} = \theta_i\,(1-I_i) + \theta_i^{\text{bias}}\,I_i$$

where $I_i \in \{0,1\}$ is a latent indicator:

- **Component 1** ($I_i=0$, *unbiased*): $\theta_i \sim N(\mu_1, \tau^2)$
- **Component 2** ($I_i=1$, *biased*): $\theta_i^{\text{bias}}$ follows a
  **slash distribution** (heavier tails than Normal), centred at
  $\mu_2 = \mu_1 + B$.

$\mu_1$ is the quantity of interest — the bias-corrected pooled effect. $B$ is
the estimated systematic inflation across biased studies.

## Analogy to the TAAR1 data

In our dataset the animal studies play the role of "observational" evidence
and the RCTs play the role of the unbiased reference arm, directly mirroring
the Verde (2021) `ppvipd` worked example.

## PBias extension

The **PBias model** (Verde 2021, Section 3.2.6) replaces the single global
bias fraction $p_{\text{bias}}$ with a per-study logistic regression:

$$\text{logit}(P[i, 2]) = \alpha_0 + \alpha_1 x_i$$

where $x_i$ is a study-level risk-of-bias (ROB) score. A positive $\alpha_1$
means studies with higher ROB scores are more likely to be in the biased
component.

---

# Packages

```{r packages}
# Install if needed (comment out once installed):
# install.packages(c("jarbes", "rjags", "R2jags", "LearnBayes"))

library(jarbes)
library(rjags)
library(R2jags)
library(LearnBayes)
library(dplyr)
library(readr)
library(stringr)
```

---

# Data

## Animal studies: drug-level SMDs

`taar1_drug_merged.csv` contains one row per drug-level animal meta-analytic
estimate. The estimates are standardised mean differences (SMDs) on locomotor
activity, derived from a multivariate Bayesian meta-analysis of SYRCLE-assessed
animal studies (see `taar1-agonists.R`). The dataset also carries the ten
SYRCLE domain judgements used later in the PBias model section.

```{r load-animal}
taar1_animal_raw <- read_csv("taar1_drug_merged.csv")

taar1_animal <- taar1_animal_raw %>%
  rename(name = DrugName) %>%
  mutate(
    design = "Animal",
    TE     = yi,
    seTE   = sqrt(vi)
  ) %>%
  select(name, design, TE, seTE, everything())

taar1_animal %>%
  select(name, TE, seTE) %>%
  mutate(across(c(TE, seTE), \(x) round(x, 3))) %>%
  knitr::kable(caption = paste0("Animal drug-level SMDs (n = ", nrow(taar1_animal), ")"))
```

## Human RCTs: trial-level SMDs

`human_taar1.csv` contains four Phase II/III RCTs comparing a TAAR1 agonist
(ulotaront or ralmitaront) against placebo in adults with acute schizophrenia.
Effect sizes are SMDs on overall symptom severity (PANSS total or equivalent),
assessed with ROB2.

```{r load-human}
taar1_human_raw <- read_csv("human_taar1.csv")

taar1_human <- taar1_human_raw %>%
  rename(name = study_name_drug) %>%
  mutate(
    design = "RCT",
    TE     = yi,
    seTE   = sqrt(vi)
  ) %>%
  select(name, design, TE, seTE, everything())

taar1_human %>%
  select(name, TE, seTE) %>%
  mutate(across(c(TE, seTE), \(x) round(x, 3))) %>%
  knitr::kable(caption = paste0("Human RCT SMDs (n = ", nrow(taar1_human), ")"))
```

## Combined dataset

```{r combine-data}
taar1_combined <- bind_rows(taar1_animal, taar1_human)

taar1_combined %>%
  group_by(design) %>%
  summarise(
    n       = n(),
    min_TE  = round(min(TE), 3),
    mean_TE = round(mean(TE), 3),
    max_TE  = round(max(TE), 3),
    .groups = "drop"
  ) %>%
  knitr::kable(caption = "Effect size range by evidence type (SMD)")
```

The animal studies tend to show much larger effects than the RCTs
(mean animal SMD ≈ 1.1 vs mean RCT SMD ≈ 0.15), which is precisely the kind
of systematic discrepancy the BC model is designed to quantify and correct.

---

# BUGS model: Bayesian random-effects meta-analysis {#re-model}

The Bayesian RE model used for the separate RCT-only and animal-only analyses
follows Verde (2021, Supplementary). It is a standard normal-normal hierarchical
model with a **half-Cauchy prior on the between-study SD** $\tau$, implemented
via `dscaled.gamma(scale, df)` in JAGS.

**Likelihood:**
$$y_i \mid \theta_i \sim N(\theta_i,\; \sigma_i^2)$$

**Random effects:**
$$\theta_i \mid \mu, \tau \sim N(\mu, \tau^2)$$

**Priors:**
$$\mu \sim N(0, 1000), \quad \tau \sim \text{half-Cauchy}(0.5)$$

The half-Cauchy scale of 0.5 SMD units is the Verde (2021) default and is
weakly informative — it places most prior mass on heterogeneity below 1 SMD
while allowing larger values.

```{r write-re-model}
cat(
  "model
{
  for (i in 1:N) {
    y[i]      ~ dnorm(theta[i], pre.y[i])
    pre.y[i] <- pow(se.y[i], -2)
    theta[i]  ~ dnorm(mu, prec.tau)
  }

  tau        <- 1 / sqrt(prec.tau)
  prec.tau   ~ dscaled.gamma(scale.sigma.between, df.scale.between)

  mu         ~ dnorm(0.0, 0.001)

  theta.new  ~ dnorm(mu, prec.tau)
}",
  file = "meta_bayes.bugs"
)
```

`theta.new` is the posterior predictive effect for a hypothetical new study —
a useful shrinkage estimator for assessing translational generalisability.

## Shared MCMC settings

```{r mcmc-settings}
n.chains   <- 4
n.iter     <- 50000
n.thin     <- 1
n.burnin   <- 20000

scale.sigma.between <- 0.5   # half-Cauchy scale for tau (Verde 2021 default)
df.scale.between    <- 1

par.re <- c("theta", "mu", "theta.new", "tau")

set.seed(2026)
```

---

# Separate RE models: RCT-only and animal-only {#separate-re}

## RCT-only model

```{r fit-rct, cache=TRUE}
y.rct    <- taar1_human$TE
se.y.rct <- taar1_human$seTE
N.rct    <- length(y.rct)

data.rct <- list(
  y                   = y.rct,
  se.y                = se.y.rct,
  N                   = N.rct,
  scale.sigma.between = scale.sigma.between,
  df.scale.between    = df.scale.between
)

m.rct <- R2jags::jags(
  data               = data.rct,
  inits              = NULL,
  parameters.to.save = par.re,
  model.file         = "meta_bayes.bugs",
  n.chains           = n.chains,
  n.iter             = n.iter,
  n.thin             = n.thin,
  n.burnin           = n.burnin,
  DIC                = TRUE
)

```

<details>
<summary>Show full JAGS output: RCT-only model</summary>
```{r print-rct, echo=FALSE, cache=TRUE, cache.extra=m.rct$BUGSoutput$DIC}
print(m.rct)
```
</details>

```{r rct-summary}
attach.jags(m.rct, overwrite = TRUE)
mu.rct <- mu

post_summary(mu.rct, "mu – pooled SMD") %>%
  knitr::kable(caption = "RCT-only model: posterior summary")
```

With only four trials the posterior is wide, but the mean (~0.15) is
consistent with the frequentist PANSS meta-analysis from the GALENOS review.

## Animal-only model

```{r fit-animal, cache=TRUE}
y.animal    <- taar1_animal$TE
se.y.animal <- taar1_animal$seTE
N.animal    <- length(y.animal)

data.animal <- list(
  y                   = y.animal,
  se.y                = se.y.animal,
  N                   = N.animal,
  scale.sigma.between = scale.sigma.between,
  df.scale.between    = df.scale.between
)

m.animal <- R2jags::jags(
  data               = data.animal,
  inits              = NULL,
  parameters.to.save = par.re,
  model.file         = "meta_bayes.bugs",
  n.chains           = n.chains,
  n.iter             = n.iter,
  n.thin             = n.thin,
  n.burnin           = n.burnin,
  DIC                = TRUE
)

```

<details>
<summary>Show full JAGS output: animal-only model</summary>
```{r print-animal, echo=FALSE, cache=TRUE, cache.extra=m.animal$BUGSoutput$DIC}
print(m.animal)
```
</details>

```{r animal-summary}
attach.jags(m.animal, overwrite = TRUE)
mu.animal <- mu

post_summary(mu.animal, "mu – pooled SMD") %>%
  knitr::kable(caption = "Animal-only model: posterior summary")
```

## Comparing the two separate models

```{r compare-separate}
rct_cri    <- unname(quantile(mu.rct,    probs = c(0.025, 0.975)))
animal_cri <- unname(quantile(mu.animal, probs = c(0.025, 0.975)))

mu.rct_resamp   <- sample(mu.rct, length(mu.animal), replace = TRUE)
p_animal_gt_rct <- mean(mu.animal > mu.rct_resamp)

data.frame(
  Model            = c("RCT-only", "Animal-only"),
  `Mean [95% CrI]` = c(
    sprintf("%.3f [%.3f, %.3f]", mean(mu.rct),    rct_cri[1],    rct_cri[2]),
    sprintf("%.3f [%.3f, %.3f]", mean(mu.animal), animal_cri[1], animal_cri[2])
  ),
  `CrI width` = round(c(diff(rct_cri), diff(animal_cri)), 3),
  check.names = FALSE
) %>%
  knitr::kable(
    caption = paste0(
      "Separate RE models — P(animal mu > RCT mu) = ",
      round(p_animal_gt_rct, 3)
    )
  )
```

The posterior probability that the animal pooled effect exceeds the RCT
pooled effect is very high (~1.0), consistent with the known translational
inflation pattern. The BC model will estimate this gap as $B$.

---

# BUGS model: Bias-Corrected (BC) model {#bc-model}

The BC model (Verde 2021, Section 2.3) extends the RE model by introducing a
**latent allocation indicator** $T_i \in \{1, 2\}$ for each study. Studies
allocated to component 2 (biased) have their effects drawn from a slash
distribution centred at $\mu_2 = \mu_1 + B$, where $B > 0$ represents the
systematic upward bias.

**Key structural features:**

- **Slash distribution for biased component**: heavier tails than Normal,
  implemented as $\text{Normal}(\mu_2,\; \tau^2 / w_i)$ where
  $w_i \sim \text{Beta}(0.5, 1)$. This prevents a few animal studies with
  very large effects from pulling the bias estimate.
- **Identifiability constraint**: studies are sorted by ascending TE before
  fitting. The study with the smallest TE is fixed to component 1 (unbiased
  anchor) and the study with the largest TE to component 2 (biased anchor).
  This breaks label-switching in the JAGS `dcat` sampler.
- **$B_{\text{max}} = 3$ SMD units**: the observed gap between the animal
  ($\approx 1.1$) and RCT ($\approx 0.15$) pooled means is ~0.96 SMD, so
  $B_{\text{max}} = 3$ is deliberately conservative and avoids truncating
  the prior near the observed difference.

```{r write-bc-model}
cat(
  "model
{
  for (i in 1:N) {
    y[i]          ~ dnorm(theta.bc[i], pre.y[i])
    pre.y[i]      <- pow(se.y[i], -2)

    theta.bc[i]   <- theta[i] * (1 - I[i]) + theta.bias[i] * I[i]
    I[i]          <- T[i] - 1

    theta[i]      ~ dnorm(mu[1], prec.tau[i])
    theta.bias[i] ~ dnorm(mu[2], prec.tau[i])

    T[i]          ~ dcat(p.bias[1:2])

    prec.tau[i]   <- inv.var[T[i]] * w[T[i], i]
    w[1, i]       <- 1
    w[2, i]       ~ dbeta(nu, 1)
  }

  nu <- 1/2

  p.bias[2] ~ dbeta(alpha.bias, beta.bias)
  p.bias[1] <- 1 - p.bias[2]

  tau          <- 1 / sqrt(inv.var[1])
  inv.var[1]   ~ dscaled.gamma(scale.sigma.between, df.scale.between)
  inv.var[2]   <- inv.var[1]

  mu[1] ~ dnorm(0.0, 0.01)
  B     ~ dunif(0, B.max)
  mu[2] <- mu[1] + B
}",
  file = "BC.bugs"
)
```

## Prepare the combined sorted dataset

```{r prepare-bc-data}
taar1_sorted <- taar1_combined %>% arrange(TE)

y.bc    <- taar1_sorted$TE
se.y.bc <- taar1_sorted$seTE
N.bc    <- nrow(taar1_sorted)

# Fixed anchors: smallest TE -> unbiased (1), largest TE -> biased (2)
T.init        <- rep(NA, N.bc)
T.init[1]     <- 1
T.init[N.bc]  <- 2

B.max <- 3

bind_rows(head(taar1_sorted, 5), tail(taar1_sorted, 5)) %>%
  select(name, design, TE, seTE) %>%
  mutate(across(c(TE, seTE), \(x) round(x, 3))) %>%
  knitr::kable(caption = "Combined dataset sorted by TE (first 5 and last 5 rows)")
```

## Informative prior on the bias fraction {#bc-inf}

We follow Verde (2021) and place a Beta prior on $p_{\text{bias}}[2]$ (the
fraction of biased studies) with:

- **Median** at $(N_{\text{animal}} - 1) / N_{\text{total}}$ ≈ 16/21
- **90th percentile** at $N_{\text{animal}} / N_{\text{total}}$ ≈ 17/21

This concentrates prior mass around the fraction of animal studies while
allowing some probability mass above and below — consistent with the
expectation that most animal studies are biased but not certifying it.

```{r prior-pbias}
N.ani.bc <- nrow(taar1_animal)

median.prior <- (N.ani.bc - 1) / N.bc
percentile90 <- N.ani.bc / N.bc

beta.par   <- LearnBayes::beta.select(
  list(p = 0.50, x = median.prior),
  list(p = 0.90, x = percentile90)
)
alpha.bias <- beta.par[1]
beta.bias  <- beta.par[2]

data.frame(
  Quantity = c("Prior median", "Prior 90th percentile", "Beta alpha", "Beta beta"),
  Value    = round(c(median.prior, percentile90, alpha.bias, beta.bias), 3)
) %>%
  knitr::kable(caption = "Informative Beta prior on p.bias")
```

## Fit the BC model (informative prior)

```{r fit-bc-inf, cache=TRUE}
data.bc <- list(
  y                   = y.bc,
  se.y                = se.y.bc,
  N                   = N.bc,
  T                   = T.init,
  scale.sigma.between = scale.sigma.between,
  df.scale.between    = df.scale.between,
  alpha.bias          = alpha.bias,
  beta.bias           = beta.bias,
  B.max               = B.max
)

par.bc <- c(
  "theta.bc", "theta", "theta.bias",
  "mu", "tau", "p.bias", "I", "w", "B"
)

set.seed(2026)

bc.taar1 <- R2jags::jags(
  data               = data.bc,
  inits              = NULL,
  parameters.to.save = par.bc,
  model.file         = "BC.bugs",
  n.chains           = n.chains,
  n.iter             = 50000,
  n.thin             = 4,
  n.burnin           = 20000,
  DIC                = TRUE
)

```

<details>
<summary>Show full JAGS output: BC model (informative prior)</summary>
```{r print-bc-inf, echo=FALSE, cache=TRUE, cache.extra=bc.taar1$BUGSoutput$DIC}
print(bc.taar1)
```
</details>

## BC model results (informative prior)

```{r bc-inf-summary}
attach.jags(bc.taar1, overwrite = TRUE)

mu.bc.corrected <- mu[, 1]
mu.bc.biased    <- mu[, 2]
B.posterior     <- B
p.bias.post     <- p.bias[, 2]
eff.unbiased    <- mean((1 - p.bias.post) * N.bc)

bind_rows(
  post_summary(mu.bc.corrected, "mu[1] — bias-corrected pooled SMD"),
  post_summary(mu.bc.biased,    "mu[2] — biased component mean"),
  post_summary(B.posterior,     "B — systematic bias (mu[2] - mu[1])"),
  post_summary(p.bias.post,     "p.bias — fraction of biased studies")
) %>%
  knitr::kable(
    caption = paste0(
      "BC model (informative prior) — effective unbiased studies: ",
      round(eff.unbiased, 1), " of ", N.bc
    )
  )
```

## Study-level posterior probability of being biased

```{r bc-study-bias}
I.post <- apply(I, 2, mean)

study_bias_summary <- taar1_sorted %>%
  select(name, design, TE, seTE) %>%
  mutate(
    TE       = round(TE, 3),
    seTE     = round(seTE, 3),
    P_biased = round(I.post, 3)
  ) %>%
  arrange(desc(P_biased))

study_bias_summary %>%
  head(10) %>%
  knitr::kable(caption = "Study-level P(biased): top 10 (BC, informative prior)")
```

Studies with $P(\text{biased}) \approx 1$ are those the model is near-certain
belong to the inflated component. Expect most animal studies to cluster here.
RCTs with low ROB2 ratings should have $P(\text{biased})$ close to 0.

```{r save-bc-inf}
mu.bc.corrected.inf <- mu.bc.corrected
B.posterior.inf     <- B.posterior
p.bias.post.inf     <- p.bias.post
```

---

# BC model with non-informative prior on p.bias {#bc-noninf}

Verde (2021, Section 3.2.2) recommends a sensitivity check using a **Uniform
prior** on the bias fraction ($\alpha = \beta = 1$). This removes the
assumption that ~17/21 studies are biased and tests whether the *data alone*
can separate the two components purely from the observed effect-size
distribution.

If $\mu_1$ from this model agrees with the informative-prior result, the
correction is robust to the prior specification.

```{r fit-bc-noninf, cache=TRUE}
data.bc.noninf            <- data.bc
data.bc.noninf$alpha.bias <- 1
data.bc.noninf$beta.bias  <- 1

set.seed(2026)

bc.taar1.noninf <- R2jags::jags(
  data               = data.bc.noninf,
  inits              = NULL,
  parameters.to.save = par.bc,
  model.file         = "BC.bugs",
  n.chains           = n.chains,
  n.iter             = 50000,
  n.thin             = 4,
  n.burnin           = 20000,
  DIC                = TRUE
)

```

<details>
<summary>Show full JAGS output: BC model (non-informative prior)</summary>
```{r print-bc-noninf, echo=FALSE, cache=TRUE, cache.extra=bc.taar1.noninf$BUGSoutput$DIC}
print(bc.taar1.noninf)
```
</details>

```{r bc-noninf-summary}
attach.jags(bc.taar1.noninf, overwrite = TRUE)

mu.bc.noninf  <- mu[, 1]
B.noninf      <- B
p.bias.noninf <- p.bias[, 2]

# Verde (2021) diagnostic: P(B > 2*tau) tests whether a meaningful bias
# component is supported by the data
tau.noninf    <- tau
p.bias.exists <- mean(B.noninf > 2 * mean(tau.noninf))

bind_rows(
  post_summary(mu.bc.noninf,  "mu[1] — bias-corrected pooled SMD"),
  post_summary(B.noninf,      "B — systematic bias"),
  post_summary(p.bias.noninf, "p.bias — fraction of biased studies")
) %>%
  knitr::kable(
    caption = paste0(
      "BC model (non-informative prior) — P(B > 2τ) = ",
      round(p.bias.exists, 3)
    )
  )
```

$P(B > 2\tau)$ is the Verde (2021) diagnostic for whether a meaningful bias
component is present in the data. Values close to 1 indicate the two
components are well separated.

```{r save-bc-noninf}
mu.bc.noninf.saved  <- mu.bc.noninf
B.noninf.saved      <- B.noninf
p.bias.noninf.saved <- p.bias.noninf
```

---

# PBias model: per-study bias probability via ROB score {#pbias-rob}

## Deriving the ROB covariate

The PBias model requires a scalar risk-of-bias covariate $x_i \in [0,1]$ for
each study, where higher values indicate worse quality. The two evidence arms
use different ROB tools so the score is derived separately and then merged.

### Animal studies (SYRCLE 10-domain tool)

Domain values: **Yes** = low risk (0), **Unclear** = moderate (0.5),
**No** = high risk (1). The ROB score is the mean across all 10 domains.

```{r rob-animal}
animal_rob_cols <- c(
  "D1 Allocation sequence", "D2 Baseline similarity",
  "D3 Allocation concealment", "D4 Random housing",
  "D5 Caregivers blinded", "D6 Random selection for outcome assessment",
  "D7 Blinded outcome assessor", "D8 Incomplete data reporting addressed",
  "D9 Free from selective outcome reporting", "D10 Free from other risks of bias"
)

animal_rob_scores <- taar1_animal %>%
  select(name, all_of(animal_rob_cols)) %>%
  rowwise() %>%
  mutate(
    rob_score = mean(
      dplyr::case_when(
        c_across(all_of(animal_rob_cols)) == "No"      ~ 1,
        c_across(all_of(animal_rob_cols)) == "Unclear" ~ 0.5,
        c_across(all_of(animal_rob_cols)) == "Yes"     ~ 0,
        TRUE                                           ~ NA_real_
      ),
      na.rm = TRUE
    )
  ) %>%
  ungroup() %>%
  select(name, rob_score)

animal_rob_scores %>%
  mutate(rob_score = round(rob_score, 3)) %>%
  knitr::kable(caption = "SYRCLE ROB scores for animal studies")
```

### Human RCTs (Cochrane ROB2 tool)

Domain values: **Low** = low risk (0), **Some concerns** or **High** = (1).
Domains marked **NA** (not applicable) are excluded from the denominator.

```{r rob-human}
human_rob_cols <- c(
  "Bias due to randomization",
  "Bias due to deviations from intended intervention",
  "Bias due to missing data",
  "Bias due to outcome measurement",
  "Bias due to selected reported results",
  "Bias due to period and carryover effects"
)

human_rob_scores <- taar1_human %>%
  select(name, all_of(human_rob_cols)) %>%
  rowwise() %>%
  mutate(
    rob_score = {
      vals       <- c_across(all_of(human_rob_cols))
      applicable <- vals[!is.na(vals) & vals != "NA"]
      if (length(applicable) == 0) NA_real_
      else mean(applicable %in% c("Some concerns", "High"))
    }
  ) %>%
  ungroup() %>%
  select(name, rob_score)

human_rob_scores %>%
  mutate(rob_score = round(rob_score, 3)) %>%
  knitr::kable(caption = "ROB2 scores for human RCTs")
```

### Join ROB scores into the sorted dataset

```{r merge-rob}
rob_lookup <- bind_rows(animal_rob_scores, human_rob_scores)

taar1_sorted <- taar1_sorted %>%
  select(-any_of("rob_score")) %>%
  left_join(rob_lookup, by = "name")

taar1_sorted %>%
  select(name, design, TE, rob_score) %>%
  arrange(design, rob_score) %>%
  mutate(TE = round(TE, 3), rob_score = round(rob_score, 3)) %>%
  knitr::kable(caption = "ROB scores by study (sorted by design then score)")

x.rob <- taar1_sorted$rob_score
```

## BUGS model: PBias

The structure is identical to the BC model except that the global
$p_{\text{bias}}[2] \sim \text{Beta}(\cdot)$ is replaced by a per-study
logistic regression on the ROB score:

$$\text{logit}(P[i, 2]) = \alpha_0 + \alpha_1 x_i$$

Priors on the logistic coefficients follow Verde (2021): $\text{Normal}(0,
\text{precision} = 0.1)$, giving $\text{SD} \approx 3.16$ on the logit scale —
weakly informative but allowing the log-odds to move freely across the
covariate range.

```{r write-pbias-model}
cat(
  "model
{
  for (i in 1:N) {
    y[i]        ~ dnorm(theta.bc[i], pre.y[i])
    pre.y[i]   <- pow(se.y[i], -2)

    theta.bc[i] <- theta[i] * (1 - I[i]) + theta.bias[i] * I[i]
    I[i]        <- T[i] - 1

    theta[i]      ~ dnorm(mu[1], prec.tau[i])
    theta.bias[i] ~ dnorm(mu[2], prec.tau[i])

    T[i] ~ dcat(P[i, 1:2])

    logit(P[i, 2]) <- alpha.0 + alpha.1 * x[i]
    P[i, 1]        <- 1 - P[i, 2]

    prec.tau[i] <- inv.var[T[i]] * w[T[i], i]
    w[1, i]     <- 1
    w[2, i]     ~ dbeta(nu, 1)
  }

  nu <- 1/2

  alpha.0 ~ dnorm(prior.mean.alpha.0, prior.scale.alpha.0)
  alpha.1 ~ dnorm(prior.mean.alpha.1, prior.scale.alpha.1)

  tau        <- 1 / sqrt(inv.var[1])
  inv.var[1] ~ dscaled.gamma(scale.sigma.between, df.scale.between)
  inv.var[2] <- inv.var[1]

  mu[1] ~ dnorm(0.0, 0.01)
  B     ~ dunif(0, B.max)
  mu[2] <- mu[1] + B
}",
  file = "PBias.bugs"
)
```

## Fit the PBias model

```{r fit-pbias, cache=TRUE}
data.pbias <- list(
  y                   = y.bc,
  se.y                = se.y.bc,
  x                   = x.rob,
  N                   = N.bc,
  T                   = T.init,
  scale.sigma.between = scale.sigma.between,
  df.scale.between    = df.scale.between,
  B.max               = B.max,
  prior.mean.alpha.0  = 0,
  prior.scale.alpha.0 = 0.1,   # precision = 0.1  =>  SD ~ 3.16
  prior.mean.alpha.1  = 0,
  prior.scale.alpha.1 = 0.1
)

par.pbias <- c(
  "theta.bc", "mu", "tau",
  "P", "T", "w", "B",
  "alpha.0", "alpha.1"
)

set.seed(2026)

PBias.taar1 <- R2jags::jags(
  data               = data.pbias,
  inits              = NULL,
  parameters.to.save = par.pbias,
  model.file         = "PBias.bugs",
  n.chains           = n.chains,
  n.iter             = 50000,
  n.thin             = 4,
  n.burnin           = 20000,
  DIC                = TRUE
)

```

<details>
<summary>Show full JAGS output: PBias model</summary>
```{r print-pbias, echo=FALSE, cache=TRUE, cache.extra=PBias.taar1$BUGSoutput$DIC}
print(PBias.taar1)
```
</details>

## PBias model results

```{r pbias-summary}
attach.jags(PBias.taar1, overwrite = TRUE)

mu.pbias     <- mu[, 1]
B.pbias      <- B
alpha.0.post <- alpha.0
alpha.1.post <- alpha.1

p_alpha1_pos <- mean(alpha.1.post > 0)

bind_rows(
  post_summary(mu.pbias,     "mu[1] — bias-corrected pooled SMD"),
  post_summary(B.pbias,      "B — systematic bias"),
  post_summary(alpha.0.post, "alpha.0 — logistic intercept"),
  post_summary(alpha.1.post, "alpha.1 — logistic slope on ROB score")
) %>%
  knitr::kable(
    caption = paste0(
      "PBias model — P(α₁ > 0) = ", round(p_alpha1_pos, 3),
      "  [positive = higher ROB score predicts higher P(biased)]"
    )
  )
```

## Per-study P(biased) from PBias model

```{r pbias-study-probs}
P.post.mean <- apply(P[, , 2], 2, mean)

study_pbias_summary <- taar1_sorted %>%
  select(name, design, TE, rob_score) %>%
  mutate(
    TE             = round(TE, 3),
    rob_score      = round(rob_score, 3),
    P_biased_PBias = round(P.post.mean, 3)
  ) %>%
  arrange(desc(P_biased_PBias))

study_pbias_summary %>%
  knitr::kable(caption = "Per-study P(biased) from PBias model")
```

---

# Final comparison across all models {#final-comparison}

```{r final-comparison}
width_rct <- diff(quantile(mu.rct,              probs = c(0.025, 0.975)))
width_bc  <- diff(quantile(mu.bc.corrected.inf, probs = c(0.025, 0.975)))

bind_rows(
  post_summary(mu.rct,              "1. RCT-only RE"),
  post_summary(mu.animal,           "2. Animal-only RE"),
  post_summary(mu.bc.corrected.inf, "3. BC (informative prior)"),
  post_summary(mu.bc.noninf.saved,  "4. BC (non-informative prior)"),
  post_summary(mu.pbias,            "5. PBias (ROB covariate)")
) %>%
  knitr::kable(caption = "Bias-corrected pooled SMD across all models")
```

```{r final-comparison-bias}
bind_rows(
  post_summary(B.posterior.inf, "BC (informative prior)"),
  post_summary(B.noninf.saved,  "BC (non-informative prior)"),
  post_summary(B.pbias,         "PBias (ROB covariate)")
) %>%
  knitr::kable(caption = "Estimated systematic bias B across BC models")
```

The BC model narrows the 95% CrI on the pooled SMD from
`r round(width_rct, 3)` (RCT-only) to `r round(width_bc, 3)` (BC $\mu_1$),
a `r round(width_rct / width_bc, 1)`× gain in precision from formally
borrowing the animal evidence while correcting for its bias.

## Interpretation

The two tables above allow us to answer three questions simultaneously:

1. **Does the BC model recover the RCT estimate after correction?**
   $\mu_1$ from the BC models should be close to the RCT-only pooled effect.
   Agreement across informative and non-informative priors strengthens
   confidence in the correction.

2. **How large is the systematic bias?**
   $B$ is the posterior estimate of how much the animal studies systematically
   over-estimate the true effect. A large $B$ with a narrow CrI indicates the
   mixture components are well separated.

3. **Does the ROB score predict bias?**
   A positive $\alpha_1$ in the PBias model with $P(\alpha_1 > 0)$ close to 1
   would confirm that studies with higher ROB scores are more likely to belong
   to the inflated component — validating the use of SYRCLE/ROB2 assessments
   as a quantitative bias covariate.

---

# Session information

```{r session-info}
sessionInfo()
```
